Thesis Title Goes Here

### 2024-05-06 introduction/intro.tex
Updates from Overleaf

In the innermost zone of the star the Sun has a temperature of about 15 MK and a density of $1.6~\times~10^{5}~km.m^{-3}$ that fuses Hydrogen into Helium mainly by p-p cycle and to some extent by CNO cycle \sr{Ref here}.

@@

Fig.~\ref{fig_solar_int} shows a schematic diagram of various layers of the Sun. Beyond the core we have the optically thick radiative zone, where the Hydrogen and Helium are completely ionized \sr{Ref. here}. The high energy $\gamma$-ray photons generated in the core, have very small mean free path in this region, as they collide numerous times eventually thermalizing becoming visible photon. Beyond the radiative zone the degree of ionization of Hydrogen and Helium changes drastically giving rise to a sharp temperature gradient. This is the convective zone where the energy from the inner layers of the Sun is transported adiabetically to the surface of the Sun. In between the radiative and the convective zone, there is a thin layer, roughly at $\sim~0.7~R_{\odot}$ known as the tacholine, which is believed to play a major role in generating Sun's magnetic field via the Dynamo process \sr{Ref. here}.

@@

The Photosphere is defined to be the layer where $\tau_{\lambda~\sim~5000}~=~1$. This is in the green part of the visible spectrum and the Sun is opaque in visible beyond this layer, hence it is called to be the surface of the Sun. This layer is 400-600 km deep and has an effective temperature of 5780 K. The magnetic filed lines arising from the tacholine penetrate the Photosphere and creates a `carpet' over the whole region \sr{(Priest 2014) ref here}. As mentioned earlier, this is the Solar surface known as the `Quite Sun'(QS from hereon). This region exhibits an average magnetic flux density of \textbf{10 -- 50~G}. The QS surface is covered with cells of roughly four size- granule, meso granules, super granules and giant cells. Some regions exhibit much stronger magnetic flux density often associated with highly twisted magnetic field structures. These magnetic features are manifested as sunspots, spicules, bright points etc.

As the density changes drastically at the Photosphere compared to the convective zone, the thermalized photons from Sun's core start free streaming again, as the mean free path also increases drastically and as a result perturbs the thermal equilibrium (TE from hereon). So, we have to define the thermodynamic quantities of the Photosphere in Local Thermal Equilibrium (LTE from hereon). As we move away from the Photosphere, because the density keeps on decreasing steadily, the LTE conditions also start deviating similarly \sr{Ref here}.

@@

The Chromosphere is a highly non-uniform dynamic layer with a thickness of 1500~--~3000~km, with increasing temperature (upto $10^{4}$ K) and decreasing number density. As seen in the fig. \ref{fig_solar_atm}, the temperature of the Chromosphere saturates before the dramatic rise in the transition region. This saturation is attributed to a steady deposition of acoustic energy by creation of shock waves \sr{Ref. here}. The Chromosphere also exhibits a sharp gradient in the plasma $\beta$ factor, non-LTE conditions, dominance of wave motions \sr{Ref. here}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[h!]
    \centering
    \includegraphics[width=0.5\linewidth]{Figures/phu_63_8_818_f2.jpg}
    \caption{A schematic diagram of the standard model of solar flare.}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

It is generally well known by now that the active region have enough magnetic free energy to power the flare and CME \citep{emslie12,ash17}. The partition of energy between the flare and associated CME is much more fuzzy. \cite{emslie12} found that the flare and CME have energies of same order of magnitude, while \cite{ash17} concluded the flare dominates the CME in-terms of the energy. However the simple question, whether the non-thermal component of the flares have enough enrgy is still not resolved as even the most recent studies contradict each other in the most puzzling fashion. Here I will discuss the contradictions arising from some of the studies\citep{stosire07,emslie12,inglis14,warmuth16a,warmuth16b,ash17}. The details of the studies can be summarized as follows:

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[h!]
    \centering
    \hline 
    \hline
    \resizebox{\textwidth}{!}{%
    \begin{tabular}{||c|c|c|c|c|c|c||}
       Study & No. of flares & GOES class range & Thermal model & Thermal spectrum & Thermal volume & Thermal losses \\
       \cite{stosire07} & 18 & A3-B7 & Isotherm. & RHESSI & TRACE & X \\
       \cite{emslie12} & 38 & C5-X28 & Isotherm. & RHESSI & RHESSI & Rad. \\
       \cite{inglis14} & 10 & B3-B9 & Multitherm. & RHESSI+AIA & RHESSI & Rad. \\
       \cite{warmuth16a,warmuth16b} & 24 & C3-X17 & Isotherm. & RHESSI+GOES & RHESSI & Rad.,Cond. \\
       \cite{ash17} & 188 & M1-X7 & Multitherm. & AIA & AIA & Rad. \\
       \hline
    \end{tabular}}
    \caption{The details of the studies.}
    \label{tab1}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

@@

>>>>

In the innermost zone of the star the Sun has a temperature of about 15 MK and a density of $1.6~\times~10^{5}~km.m^{-3}$ that fuses Hydrogen into Helium mainly by p-p cycle \citep{bethe38} and to some extent by CNO cycle \citep{bethe39}.

@@

Fig.~\ref{fig_solar_int} shows a schematic diagram of various layers of the Sun. Beyond the core we have the optically thick radiative zone, where the Hydrogen and Helium are completely ionized. The high energy $\gamma$-ray photons generated in the core, have very small mean free path in this region, as they collide numerous times eventually thermalizing becoming visible photon. Beyond the radiative zone the degree of ionization of Hydrogen and Helium changes drastically giving rise to a sharp temperature gradient. This is the convective zone where the energy from the inner layers of the Sun is transported adiabetically to the surface of the Sun. In between the radiative and the convective zone, there is a thin layer, roughly at $\sim~0.7~R_{\odot}$ known as the tacholine, which is believed to play a major role in generating Sun's magnetic field via the Dynamo process \citep{corbard01}.

@@

The Photosphere is defined to be the layer where $\tau_{\lambda~\sim~5000}~=~1$. This is in the green part of the visible spectrum and the Sun is opaque in visible beyond this layer, hence it is called to be the surface of the Sun. This layer is 400-600 km deep and has an effective temperature of 5780 K. The magnetic filed lines arising from the tacholine penetrate the Photosphere and creates a `carpet' over the whole region \citep{priest14}. As mentioned earlier, this is the Solar surface known as the `Quite Sun'(QS). This region exhibits an average magnetic flux density of \textbf{10 -- 50~G}. The QS surface is covered with cells of roughly four size- granule, meso granules, super granules and giant cells. Some regions exhibit much stronger magnetic flux density often associated with highly twisted magnetic field structures. These magnetic features are manifested as sunspots, spicules, bright points etc.

As the density changes drastically at the Photosphere compared to the convective zone, the thermalized photons from Sun's core start free streaming again, as the mean free path also increases drastically and as a result perturbs the thermal equilibrium (TE from hereon). So, we have to define the thermodynamic quantities of the Photosphere in Local Thermal Equilibrium (LTE from hereon). As we move away from the Photosphere, because the density keeps on decreasing steadily, the LTE conditions also start deviating similarly \citep{philips08}.

@@

The Chromosphere is a highly non-uniform dynamic layer with a thickness of 1500~--~3000~km, with increasing temperature (upto $10^{4}$ K) and decreasing number density. As seen in the fig. \ref{fig_solar_atm}, the temperature of the Chromosphere saturates before the dramatic rise in the transition region. This saturation is attributed to a steady deposition of acoustic energy by creation of shock waves \citep{carlsson07}. The Chromosphere also exhibits a sharp gradient in the plasma $\beta$ factor, non-LTE conditions, dominance of wave motions \sr{Ref. here}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[h!]
    \centering
    \includegraphics[width=0.5\linewidth]{Figures/phu_63_8_818_f2.jpg}
    \caption{A schematic diagram of the standard model of solar flare (adapted from \cite{lysenko20}).}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

It is generally well known by now that the active region have enough magnetic free energy to power the flare and CME \citep{emslie12,ash17}. The partition of energy between the flare and associated CME is much more fuzzy. \cite{emslie12} found that the flare and CME have energies of same order of magnitude, while \cite{ash17} concluded the flare dominates the CME in-terms of the energy. However the simple question, whether the non-thermal component of the flares have enough enrgy is still not resolved as even the most recent studies contradict each other in the most puzzling fashion. \cite{warmuth20} discussed the contradictions arising from some of the studies\citep{stosire07,emslie12,inglis14,warmuth16a,warmuth16b,ash17}. The details of the studies can be summarized as follows:

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[h!]
    \centering
    \hline 
    \hline
    \resizebox{\textwidth}{!}{%
    \begin{tabular}{||c|c|c|c|c|c|c||}
       Study & No. of flares & GOES class range & Thermal model & Thermal spectrum & Thermal volume & Thermal losses \\
       \cite{stosire07} (S07 from hereon) & 18 & A3-B7 & Isotherm. & RHESSI & TRACE & X \\
       \cite{emslie12} (E12 from hereon) & 38 & C5-X28 & Isotherm. & RHESSI & RHESSI & Rad. \\
       \cite{inglis14} (IC14 from hereon) & 10 & B3-B9 & Multitherm. & RHESSI+AIA & RHESSI & Rad. \\
       \cite{warmuth16a,warmuth16b} (WM16 from here on) & 24 & C3-X17 & Isotherm. & RHESSI+GOES & RHESSI & Rad.,Cond. \\
       \cite{ash17} (A17 from here on) & 188 & M1-X7 & Multitherm. & AIA & AIA & Rad. \\
       \hline
    \end{tabular}}
    \caption{The details of the studies.}
    \label{tab1}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The bolometric energy serves as a representation of the overall energy released during solar flares. Regardless of how energy is released in a solar flare, whether through direct plasma heating, rapid bulk flows, or nonthermal particle processes—the ultimate outcome is thermalization and subsequent radiation across the electromagnetic spectrum. Given that bolometric energy encompasses the entire spectrum, it is anticipated to reflect the total energy originally released. It's important to note that this specifically pertains to the energy liberated within the flare and does not encompass the energy associated with a concurrent coronal mass ejection (CME) or filament eruption. So, \cite{warmuth20} used the bolometric energy released in solar flares, to compare various studies on equal footing. Here we will discuss some of the main issues \cite{warmuth20} identifies comparing these studies.

The calculated thermal energy of all studies show an excellent correlation with the {\it GOES} class. Fig.~\ref{fig:goes-therm} left panel shows the peak thermal energy as a function of peak {\it GOES} flux for all the flares from the five studies. EM12 and WM16 are consistently one order of magnitude lower than the bolometric energy. Interpolation of E12 and WM16 are about half an order of magnitude lower compared to IC14, while S07 is very consistent with the extrapolation. The energies of A17 is about an order of magnitude higher compared to other studies, and even higher than the extrapolated bolometric energy for flares of same class from \cite{kretzschmar11} and E12. There can be several reason of these differences between the studies arising.

A possible reason can be attributed to different thermal models used to model the thermal component of the flares. {\it RHESSI} usually yields a higher temperature and lower emission measure compared to {\it GOES} due to the highly multi thermal nature of the flaring plasma \citep{bataglia05,ryan14,warmuth16a}. Although the {\it GOES} temperature is higher by a factor of 1.4, it is not dependant on the flare class. Along with that, \cite{warmuth20} also investigated whether assuming a isothermal or multithermal model introduces any differences in the thermal energy estimates. They reported no significant differences due to differences of models (refer to Fig. 2 and the corresponding discussion in \cite{warmuth20}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[ht!]
    \centering
    \includegraphics[width=0.45\textwidth,trim={0cm 0cm 0cm 0.02cm},clip]{goes_flux_therm.jpg}
    \includegraphics[width=0.45\textwidth,trim={0cm 0cm 0cm 0.02cm},clip]{goes_flux_vol.jpg}
    \caption{{\it GOES} peak thermal energy (left panel) and volume of the thermal plasma (right panel) as a function of the peak {\it GOES} flux for all the flares from five studies (figure credit: \cite{warmuth20}). The C values show the linear correlation coefficient for various studies.}
    \label{fig:goes-therm}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The thermal energy estimated usually depends on the source volume, $U_{Th}\sim V$. Apart from S07 and A17, all other studies used {\it RHESSI} imaging to estimate the flare volume with a filling factor $f=1$. S07 determined the volume by employing a semicircular loop model, utilizing the cross-sectional area and loop length derived from the observed areas and separations of foot point brightening of 1600 {\AA} by {\it TRACE}. In contrast, A17 estimated the volume based on the flare area exceeding a certain threshold in the emission measure obtained from the spatial synthesis DEM method. Fig.~\ref{fig:goes-therm} right panel shows the volumes used for various studies to estimate the thermal energy as a function of the peak {\it GOES} flux. It is worth noting that the volume estimates from E12 and WM16 align closely, as do those from A+17, despite employing entirely different methodologies. In contrast, the micro flare volumes reported by IC14 are notably larger, ranging from one to two orders of magnitude beyond what would be anticipated based on the findings of the other four studies. Intriguingly, this discrepancy aligns with the additional volume required to account for the higher thermal energies observed in IC14. The uncertainty in detecting volume can be due to the CLEAN algorithm systematically overestimates the source size \citep{warmuth13a}. There has also been studies demonstrating that {\it RHESSI} has difficulties in resolving small sources \citep{dennis09,warmuth13b}, that would be applicable for small thermal sources in the micro flares. This illustrates the requirement for better volume estimation for better estimates of the thermal energy of flares.

We can conclude a few more puzzling questions from these studies as a whole:

%%%%%%%%%
\begin{itemize}
    \item The primary challenge in estimating the thermal energy of the hot plasma stems from determining the distribution of the Differential Emission Measure (DEM), and the radiating plasma volume.
    \item The dissipation of energy from the hot plasma is substantial. Although there is widespread agreement regarding radiative losses across studies, the extent of conductive losses disagree significantly between the studies.
    \item The non-thermal energy in injected electrons is strongly dependent on the poorly constrained low-energy cutoff.
    \item The thermal and non-thermal energy partition changes with flare class. In smaller flares, there appears to be a deficiency of energetic electrons, whereas in larger flares, the injected non-thermal energy seems to sufficient in powering the thermal component. Does this signify the existance of an additional third heating source?
\end{itemize}
%%%%%%%%%

@@

The main motivation of this thesis is to add understanding of effects of flares on the surrounding environment in lower and upper solar atmosphere, to answer some of the questions posed above. We have used imaging and spectroscopic observations from multiple space based observatories ({\it e.g.} AIA, HMI, IRIS, XRT, etc.). This research contributes to creating a perspective of flare energetics and its effect on the surrounding plasma environment from Chromosphere to Corona.

The solar flares are the strongest magnetic event in our solar system connecting various layers of the solar atmosphere ({\it e.g.} from Photosphere to Corona), coupled via mass and energy transport. While the Coronal manifestation of solar flares have been studied extensively with several instruments, there are only a few imaging and spectroscopic instruments observing the Photosphere to transition region. These include the {\it IRIS} \citep{iris}, AIA 1600, 1700, and 304 {\AA} \citep{aia} and some the filters of {\it Hinode}/SOT \citep{sot}. In addition to these, we also have had balloon experiment {\it SUNRISE} \citep{sunrise1,sunrise2}, which provided high resolution imaging of the transition regiopn, from the SuFI instrument \citep{sufi}.

This illustrates the necessity of more observatories operating at the transition region wavelengths for more comprehensive understanding about how the solar flares manifest and affect the dynamics and heating of the lower solar atmosphere. In this regard, {\suit} onboard Aditya-L1 will play a crucial role in probing the lower solar atmosphere. The various imaging channels of {\suit} would observe the manifestation of flares in transition region at various heights (NB3 $\implies$ \ion{Mg}{2} k 279.6 nm, NB4 $\implies$ \ion{Mg}{2} h 280.3 nm, NB8 $\implies$ \ion{Ca}{2} h 396.8 nm), further lower in the transition region (NB2 $\implies$ blue wing of Mg window, NB5 $\implies$ red wing of Mg window) and Sunspots and possible signature of umbral brightening due to flares at the Photosphere (NB6 300 nm and NB7 388 nm) in full disk. This would also allow us to probe the interaction of the flaring region, with other active regions on disk, giving us a comprehensive overview of the effect of solar flares and their manifestation in the lower solat atmosphere.

@@

The work within this thesis spans across various layers of the solar atmosphere to understand how solar flare couples these layers. It is carried out partly, as various preparatory analysis for {\suit}, forward modelling {\suit} observations, designing the stellar calibration scheme for {\suit}, analysis of existing transition region observations in anticipation of {\suit} observations, analysis of coronal observations of solar flares, and analysis of first solar flare observations by {\suit}. Here we will briefly discuss the structure of the rest of the thesis:

%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{itemize}
    \item In chapter \ref{c:chap2}, we review various space based missions available and used in this thesis. We also discuss various analysis techniques used in this thesis. We also motivate the importance of {\suit} and its main science goals. We also briefly discuss the specific questions regarding solar flares we can address with {\suit}.
    \item In chapter \ref{c:chap3}, we discussed some of the initial preparatory analysis we did for {\suit}, including the throughput model for {\suit}, how the filters mounted on {\suit} were chosen, the analysis of spacecraft jitter simulation and what effect it would have on the imaging.
    \item In chapter \ref{c:chap4}, we discuss the forward modelling pipeline developed for {\suit}. We use MPS-Atlas simulation and {\it IRIS} observations, with measured transmission profiles of various optical components and the PSF to forward model {\suit} observations. We also discuss a preliminary version of the deconvolution algorithm that will be deployed in the data reduction pipeline.
    \item In chapter \ref{c:chap5}, we describe the stellar calibration scheme and the selection criteria for the standard star target. We also use a sun as a star spectra to evaluate the calibration scheme.
    \item In chapter \ref{c:chap6}, we use existing {\it IRIS} observations of three solar flares to probe how the lower solar atmosphere is affected by the flares. The ratio of the \ion{Mg}{2} k and h lines, is a proxy for the optical depth of the surrounding plasma. Hence, we use the observations of \ion{Mg}{2} lines along HMI magentic field measurements to investigate the effects of magnetic field, if any. \cite{kerr15} conducted a similar study and found signatures of localised heating in the line intensity ratio. We found, that the line intensity ratio shows a correlated increase and decrease with the flare light curve. We argue this to be signature of plasma flow associated with flare, and a change in density of the local environment. This acts as a prelude of what is achievable with {\suit}. We can use a similar method to give full disk context of optical depth, with respect to eruptive events.
\end{itemize}
%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2024-05-06 main_thesis_file.tex
local change
Updates from Overleaf

\chapter[Introduction]{Introduction}\label{c:intro}
\input{./introduction/intro}
\clearpage
%
\chapter{Existing Solar Observations and techniques}\label{c:chap2}
\chaptermark{Solar Observatories}
\input{./chapter2/ms.tex}
\clearpage
%
\chapter{Intital preparatory analysis for SUIT}\label{c:chap3}
\chaptermark{SUIT preparatory analysis}
\begin{quote}
{\em ~~~~~~~This thesis chapter originally appeared in the literature as} \\
{authors,
{\em journal reference info}}
\end{quote}
\input{./chapter3/ms.tex}
\clearpage
%
\chapter{Forward modelling SUIT observations}\label{c:chap4}
\chaptermark{SUIT forward model}
\begin{quote}
{\em ~~~~~~~This thesis chapter originally appeared in the literature as} \\
{authors,
{\em journal reference info}}
\end{quote}
\input{./chapter4/ms.tex}
%
\chapter{Stellar calibration of SUIT}\label{c:chap5}
\chaptermark{SUIT stellar calibration}
\begin{quote}
{\em ~~~~~~~This thesis chapter originally appeared in the literature as} \\
{authors,
{\em journal reference info}}
\end{quote}
\input{./chapter5/ms.tex}
%
\chapter{Effects of Solar flares on the local plasma environment from Mg II observations}\label{c:chap6}
\chaptermark{Solar Flare Energetics}
\begin{quote}
{\em ~~~~~~~This thesis chapter originally appeared in the literature as} \\
{authors,
{\em journal reference info}}
\end{quote}
\input{./chapter6/ms.tex}
\clearpage
%
\chapter{Estimating thermal and non-thermal energy partition of Solar Flares}\label{c:chap7}
\chaptermark{Solar flare energy}
\begin{quote}
{\em ~~~~~~~This thesis chapter originally appeared in the literature as} \\
{authors,
{\em journal reference info}}
\end{quote}
\input{./chapter7/ms.tex}
\clearpage
%
\chapter{First X-class flare observed from SUIT}\label{c:chap8}
\chaptermark{Solar flare energy}
\begin{quote}
{\em ~~~~~~~This thesis chapter originally appeared in the literature as} \\
{authors,
{\em journal reference info}}
\end{quote}
\input{./chapter7/ms.tex}
\clearpage
%

>>>>

\chapter[Introduction]{Introduction}\label{c:intro}
\input{./introduction/intro}
\clearpage
%
\chapter{Existing Solar Observations and Techniques: A bridge towards Solar Ultraviolet Imaging Telescope}\label{c:chap2}
\chaptermark{Solar Observatories and Analysis Techniques}
\input{./chapter2/ms.tex}
\clearpage
%
\chapter{Intital preparatory analysis for SUIT}\label{c:chap3}
\chaptermark{SUIT preparatory analysis}
\begin{quote}
{\em ~~~~~~~This thesis chapter originally appeared in the literature as} \\
{authors,
{\em journal reference info}}
\end{quote}
\input{./chapter3/ms.tex}
\clearpage
%
\chapter{Forward modelling SUIT observations}\label{c:chap4}
\chaptermark{SUIT forward model}
\begin{quote}
{\em ~~~~~~~This thesis chapter originally appeared in the literature as} \\
{authors,
{\em journal reference info}}
\end{quote}
\input{./chapter4/ms.tex}
%
\chapter{Stellar calibration of SUIT}\label{c:chap5}
\chaptermark{SUIT stellar calibration}
\begin{quote}
{\em ~~~~~~~This thesis chapter originally appeared in the literature as} \\
{authors,
{\em journal reference info}}
\end{quote}
\input{./chapter5/ms.tex}
%
\chapter{Effects of Solar flares on the local plasma environment from Mg II observations}\label{c:chap6}
\chaptermark{Solar flares Mg II}
\begin{quote}
{ ~~~~~~~This thesis chapter originally appeared in the literature as} \\
{{\em The Evolution of the Ratio of \ion{Mg}{2} Intensities During Solar Flares, Roy, S., \& Tripathi, D. 2024, ApJ, 964, 106, doi:\href{https://iopscience.iop.org/article/10.3847/1538-4357/ad2a46}{10.3847/1538-4357/ad2a46}}.}
\end{quote}
\input{./chapter6/ms.tex}
\clearpage
%
\chapter{Estimating thermal energy of two Solar Flares}\label{c:chap7}
\chaptermark{Solar flare energy}
\begin{quote}
{\em ~~~~~~~This thesis chapter originally appeared in the literature as} \\
{authors,
{\em journal reference info}}
\end{quote}
\input{./chapter7/ms.tex}
\clearpage
%
\chapter{First flares observed from {\suit}}\label{c:chap8}
\chaptermark{Solar flare energy}
\begin{quote}
{\em ~~~~~~~This thesis chapter originally appeared in the literature as} \\
{authors,
{\em journal reference info}}
\end{quote}
\input{./chapter7/ms.tex}
\clearpage
%
